VM-Centric Snapshot Deduplication with Fault Isolation for Converged Cloud Architectures

### 2014-01-31 snapshot_system/paper_background.tex
fixed

Figure~\ref{fig:collocated} illustrates a converged IaaS cloud architecture 
where
each commodity server hosts a number of virtual machines and storage of these servers
is clustered using a distributed file system~\cite{googlefs03,hdfs10,NutanixPaper}.
Each physical machine hosts multiple virtual machines.  Every virtual machine
runs its own guest operating system and accesses virtual hard disks 
stored as image files maintained by the operating system running on the
physical host.
For VM snapshot backup, file-level semantics are normally not provided.
Snapshot operations take place at the virtual device driver level, which
means no fine-grained file system metadata can be used to determine the changed data. 
%This architecture does not use a separate  backup facility for a lower cost and 
%reduces the network bandwidth consumption in transferring the raw data for backup. 
%A backup service may co-locate with other cloud services on these commodity servers and 
%stores deduplicated data  in  the storage in this cluster or in
%a separate storage cluster.

@@

In designing a VC deduplication solution, we have considered and adopted some of
the following previously-developed techniques.
\begin{itemize}
\item {\em Changed Block Tracking} --
VM snapshots can be  backed up  incrementally by identifying data segments that have
changed from the previous version of the snapshot~\cite{Clements2009,Vrable2009,TanIPDPS2011}.
Such a scheme  is  VM-centric since deduplication is localized to the snapshot
history for each VM.
%We are seeking for a tradeoff since 
%global signature comparison can deliver additional compression~\cite{Guo2011,Dong2011,extreme_binning09}.
\item {\em Stateless  Data Routing} --
One approach to scalable duplicate comparison is to use content-based hash
partitioning called stateless data routing by Dong et al.~\cite{Dong2011}.
Stateless data routing divides the deduplication work with a similarity approximation. This work 
is similar to Extreme Binning by Bhagwat et al.~\cite{extreme_binning09} and 
each request is routed  to a machine which holds
a Bloom filter  or can fetch on-disk index for additional comparison.
While this approach is VM-oblivious, it motivates us to  use  a combined signature of a dataset to narrow
VM-specific local search.
\item {\em Sampled Index for Memory Parsimony} -- 
One effective approach that reduces memory usage is 
to use a sampled index with prefetching, proposed  by Guo and Efstathopoulos\cite{Guo2011}. 
The algorithm is VM oblivious and it is not easy  to adopt for a distributed architecture. 
To use a distributed memory version of the sampled index, every deduplication request
may access a remote machine for index lookup and the overall overhead of access latency for all requests
can be significant.  
\end{itemize}

>>>>

Figure~\ref{fig:collocated} illustrates a converged IaaS cloud architecture 
where
each commodity server hosts a number of virtual machines and storage of these servers
is clustered using a distributed file system~\cite{googlefs03,hdfs10}.
Each physical machine hosts multiple virtual machines.  Every virtual machine
runs its own guest operating system and accesses virtual hard disks 
stored as image files maintained by the operating system running on the
physical host.
For VM snapshot backup, file-level semantics are normally not provided.
Snapshot operations take place at the virtual device driver level, which
means no fine-grained file system metadata can be used to determine the changed data. 
%This architecture does not use a separate  backup facility for a lower cost and 
%reduces the network bandwidth consumption in transferring the raw data for backup. 
%A backup service may co-locate with other cloud services on these commodity servers and 
%stores deduplicated data  in  the storage in this cluster or in
%a separate storage cluster.

@@

In designing a VC deduplication solution, we have considered and adopted some of
the following previously-developed techniques.
1) {\em Changed Block Tracking}. 
VM snapshots can be  backed up  incrementally by identifying data segments that have
changed from the previous version of the snapshot~\cite{Clements2009,Vrable2009,TanIPDPS2011}.
Such a scheme  is  VM-centric since deduplication is localized to the snapshot
history for each VM.
%We are seeking for a tradeoff since 
%global signature comparison can deliver additional compression~\cite{Guo2011,Dong2011,extreme_binning09}.
1) {\em Stateless  Data Routing}.
One approach to scalable duplicate comparison is to use content-based hash
partitioning called stateless data routing by Dong et al.~\cite{Dong2011}.
Stateless data routing divides the deduplication work with a similarity approximation. This work 
is similar to Extreme Binning by Bhagwat et al.~\cite{extreme_binning09} and 
each request is routed  to a machine which holds
a Bloom filter  or can fetch on-disk index for additional comparison.
While this approach is VM-oblivious, it motivates us to  use  a combined signature of a dataset to narrow
VM-specific local search.
3) {\em Sampled Index for Memory Parsimony}. 
One effective approach that reduces memory usage is 
to use a sampled index with prefetching, proposed  by Guo and Efstathopoulos\cite{Guo2011}. 
The algorithm is VM oblivious and it is not easy  to adopt for a distributed architecture. 
To use a distributed memory version of the sampled index, every deduplication request
may access a remote machine for index lookup and the overall overhead of access latency for all requests
can be significant.

### 2014-01-31 snapshot_system/paper_dedup.tex
fixed

We have studied the distribution of chunk popularity~\cite{WeiZhangIEEE} 
based on a production data trace from Alibaba with over 2500 VMs.
The distribution is Zipf-like and popular chunks dominate the distribution.
We denote $\sigma$ as the percentage of unique data chunks belonging to PDS and from the evaluation we find that
$\sigma$ with a range of 2 to 4\% can deliver a fairly competitive deduplication efficiency.
For example,
with top 2\% of the most popular chunks stored in PDS, up-to 40\% of the raw backup 
data can be eliminated by fingerprint lookup in PDS index.

@@

%\item 
\textbf{VM-centric file system block management} --
When a chunk is not detected as a duplicate to any existing chunk, this chunk will be written
to the file system. Since the backend file system typically uses a large block size such as 64MB, each VM will 
accumulate small local chunks. We manage this accumulation process using a log-structured storage scheme built
on a distributed file system discussed in Section~\ref{sect:architecture}.
%Non-PDS chunks from the same VM are stored in one append store. 
Each file system block is either dedicated to non-PDS chunks, or PDS-chunks.
A file system block for non-PDS chunks is associated with one VM and does not contain
any PDS chunks, such that our goal of fault isolation is maintained.
In addition, storing PDS chunks separately allows special replication handling for those popular shared data. 
%%% talk something about snapshot store advantage here? %%%

@@

% \begin{figure}
% \centering
%  \includegraphics[width=3in]{figures/log-log-disk}
% \caption{Duplicate frequency versus  chunk ranking in a log scale after local deduplication.}
% \label{fig:Datazipf}
% \end{figure}

>>>>

 Figure~\ref{fig:Datazipf} shows the distribution of chunk popularity for a data trace 
from Alibaba with 2500 VMs discussed in Section ~\ref{sect:evaluation}.
%the distribution of chunk popularity~\cite{WeiZhangIEEE} 
%based on a production data trace from Alibaba with over 2500 VMs.
The distribution is Zipf-like and popular chunks dominate the distribution.
We denote $\sigma$ as the percentage of unique data chunks belonging to PDS and 
from the evaluation in Section~\ref{sect:evaluation}, we find that
$\sigma$ with a range of 2 to 4\% can deliver a fairly competitive deduplication efficiency.
%For example,
%with top 2\% of the most popular chunks stored in PDS, up-to 40\% of the raw backup 
%data can be eliminated by fingerprint lookup in PDS index.

@@

%\item 
\textbf{VM-centric file system block management} --
When a chunk is not detected as a duplicate to any existing chunk, this chunk will be written
to the file system. Since the backend file system typically uses a large block size such as 64MB, each VM will 
accumulate small local chunks. 
%We manage this accumulation process using a log-structured storage scheme built
%on a distributed file system discussed in Section~\ref{sect:architecture}.
%Non-PDS chunks from the same VM are stored in one append store. 
Each file system block is either dedicated to non-PDS chunks, or PDS-chunks.
A file system block for non-PDS chunks is associated with one VM and does not contain
any PDS chunks, such that our goal of fault isolation is maintained.
In addition, storing PDS chunks separately allows special replication handling for those popular shared data. 
%%% talk something about snapshot store advantage here? %%%

@@

 \begin{figure}
 \centering
  \includegraphics[width=2.5in]{figures/log-log-disk}
 \caption{Duplicate frequency versus  chunk ranking in a log scale after local deduplication.}
 \label{fig:Datazipf}
 \end{figure}

### 2014-01-31 snapshot_system/paper_intro.tex
fixed

Commercial ``Infrastructure as a Service'' clouds (i.e. {\em public
clouds}) often make use of commodity data center components to achieve the
best possible economies of scale.  In particular, large-scale e-commerce cloud
providers such as Google and Alibaba deploy ``converged'' components that
co-locate computing and storage in each hardware module (as opposed to having
separate computing and storage ``tiers.'')  The advantage of such an approach
is that all infrastructure components are used to support paying customers --
there are no resources specifically dedicated to cloud services.
%One of emerging architectures for building cloud services is a converged
%storage architecture that leverages commodity servers with software clustered
%storage.
%The converged architecture (server plus compute in same tier) grew out of the
%Google model. 
In particular, these providers use software to
aggregate multiple direct attached low-cost disks together across
servers as a way of avoiding the relatively high cost of network attached 
storage~\cite{googlefs03,hdfs10,NutanixPaper}.
%systems. %This(e.g. ~\cite{NutanixPaper}).  Leading cloud platform companies
%such as Google, Amazon, Alibaba, Microsoft Azure  have built a converged
%compute and storage infrastructure and used a distributed file system such as
%Google file system~\cite{googlefs03,hdfs10} to glue a large number of
%commodity servers.
% with local storage into a single cluster.
In such an environment,
%That  allows the use of scalable compute and storage, without incurring the
%costs and performance limitations associated with network storage.
each physical machine runs a number of virtual machines (VMs)
%as  instances of a guest operating system 
and their virtual disks are stored as disk image files, typically
using a local (possibly shared) file system.
Frequent snapshot backup of virtual disk images can
increase the service reliability by allowing VMs to restart from their latest
snapshot in the event of a server failure.  However, because the disk
blocks that make up the VMs are often the same from snapshot to snapshot
and are shared among VMs,  
deduplication of redundant content
blocks~\cite{venti02,bottleneck08} is necessary to substantially reduce the
storage demand.
%For example, the Aliyun cloud, which is  the largest cloud service provider
%by Alibaba in China, automatically conducts  the backup of virtual disk
%images to all active users every day.  The cost of supporting a large number
%of concurrent backup streams is high because of the huge storage demand and
%the use of deduplication

>>>>

Commercial ``Infrastructure as a Service'' clouds (i.e. {\em public
clouds}) often make use of commodity data center components to achieve the
best possible economies of scale.  In particular, large-scale e-commerce cloud
providers such as Google and Alibaba deploy ``converged'' components that
co-locate computing and storage in each hardware module (as opposed to having
separate computing and storage ``tiers.'')  The advantage of such an approach
is that all infrastructure components are used to support paying customers --
there are no resources specifically dedicated to cloud services.
%One of emerging architectures for building cloud services is a converged
%storage architecture that leverages commodity servers with software clustered
%storage.
%The converged architecture (server plus compute in same tier) grew out of the
%Google model. 
In particular, these providers use software to
aggregate multiple direct attached low-cost disks together across
servers as a way of avoiding the relatively high cost of network attached 
storage~\cite{googlefs03,hdfs10}.
%systems. %This(e.g. ~\cite{NutanixPaper}).  Leading cloud platform companies
%such as Google, Amazon, Alibaba, Microsoft Azure  have built a converged
%compute and storage infrastructure and used a distributed file system such as
%Google file system~\cite{googlefs03,hdfs10} to glue a large number of
%commodity servers.
% with local storage into a single cluster.
In such an environment,
%That  allows the use of scalable compute and storage, without incurring the
%costs and performance limitations associated with network storage.
each physical machine runs a number of virtual machines (VMs)
%as  instances of a guest operating system 
and their virtual disks are stored as disk image files, typically
using a local (possibly shared) file system.
Frequent snapshot backup of virtual disk images can
increase the service reliability by allowing VMs to restart from their latest
snapshot in the event of a server failure.  However, because the disk
blocks that make up the VMs are often the same from snapshot to snapshot
and are shared among VMs,  
deduplication of redundant content
blocks~\cite{venti02,bottleneck08} is necessary to substantially reduce the
storage demand.
%For example, the Aliyun cloud, which is  the largest cloud service provider
%by Alibaba in China, automatically conducts  the backup of virtual disk
%images to all active users every day.  The cost of supporting a large number
%of concurrent backup streams is high because of the huge storage demand and
%the use of deduplication
